---
title: "Homework 3 - Problem B"
author: "Rajat Mahesh Gupta"
date: "`r Sys.Date()`"
output: pdf_document
---
Here you will develop "d,p,q,r" functions for a certain distribution
family, in the sense of e.g. Sec. 4.4.1.

- We'll call the family "accum" for "accumulate." The setting is
that of repeatedly rolling a pair of dice. The random variable
X is the number of rolls needed to achieve an accumulated total
of at least k dots. So for instance the support of X ranges
from ceiling(k/12) to ceiling(k/2). This is a one-parameter
family.
- The call forms will be
```
daccum(i,k)
paccum(i,k)
qaccum(m,k)
raccum(nreps,k)
```

- The 'd', 'p' and 'q' functions must be exact, i.e. not computed
via simulation, but a recursive call is fine. Finding 'p' and
'q' from 'd' is fine.

- For 'q', note the comment following (4.31).

```{r, cache=TRUE}
# Probability mass function for the accum distribution
daccum <- function(i, k) {
  if (i < ceiling(k / 12) || i > ceiling(k / 2)) {
    return(0)
  }
  if (i == ceiling(k / 12)) {
    return(1)
  }
  return(daccum(i - 1, k) * (1 - paccum(i - 1, k)))
}

# Cumulative distribution function for the accum distribution
paccum <- function(i, k) {
  if (i < ceiling(k / 12)) {
    return(0)
  }
  if (i >= ceiling(k / 2)) {
    return(1)
  }
  return(sum(sapply(ceiling(k / 12):i, function(j) daccum(j, k))))
}

# Quantile function for the accum distribution
qaccum <- function(m, k) {
  p <- m / 100
  for (i in ceiling(k / 12):ceiling(k / 2)) {
    if (paccum(i, k) >= p) {
      return(i)
    }
  }
  return(ceiling(k / 2))
}

# Random generation for the accum distribution
raccum <- function(nreps, k) {
  sapply(1:nreps, function(x) {
    total <- 0
    rolls <- 0
    while (total < k) {
      total <- total + sample(1:6, 1) + sample(1:6, 1)
      rolls <- rolls + 1
    }
    return(rolls)
  })
}
```

```{r, include=FALSE}
k <- 20
# Calculate probability of needing exactly 3 rolls
prob_3 <- daccum(3, 20)
cat(paste("P(X=3):", prob_3))

# Calculate probability of needing 3 or fewer rolls
cum_prob_3 <- paccum(3, 20)
cat(paste("P(X≤3):", cum_prob_3))

# Find number of rolls needed for 90th percentile
q90 <- qaccum(90, 20)
cat(paste("90th percentile:", q90))

# Generate 5 random samples
samples <- raccum(5, 20)
cat("Random samples:")
cat(samples)
```