---
title: "Practice Document"
author: "Rajat Mahesh Gupta"
date: "`r Sys.Date()`"
output: html_document
---
## Guided Example : 
If we roll three dice, what is the probability that their total is 8? We could count all the possibilities,
or we could get an approximate answer via simulation:
```{r }
# roll d dice; find P(total = k)
probtotk <- function(d,k,nreps) {
    count <- 0
    # do the experiment nreps times -- like doing nreps notebook lines
    for (rep in 1:nreps) {
        sum <- 0
        #roll d dice and find their sum 
        for (j in 1:d) sum <- sum + roll()
        if (sum == k) count <- count + 1
    }
    return (count/nreps)
}
# simulate roll of one die; the possible return values are 1,2,3,4,5,6,
# all equally likely
# roll <- function() return(sample (1:6,1))
# example
probtotk(3,8,100000)
```


The example below is the first and easiest attempt
```{r version 1 : dice total}
# roll d dice; find P(total = 8)
# d : num of dice
# k : total
# nrep : num of repetitions

pOfTotalk <- function(d, k, nrep) {
    count <- 0

    # do the experiment nrep times

    for (i in 1:nrep) {
        total <- sum(sample(1:6, d, replace = TRUE))
        if (total == k) {
            count <- count + 1
        }
    }
    return(count/nrep)
}

# simulate the roll of one dice 
pOfTotalk(3, 8, 100000000)
```

```{r version 2 : dice total}
# roll d dice; find P(total = 8)
# d : num of dice
# k : total
# nrep : num of repetitions

roll <- function(d) return(sample(1:6, d, replace = TRUE))

pOfTotalk <- function(n, k, nrep) {
    # do the experiment nrep times
    sums <- replicate(nrep, sum(roll(n)))
    return(mean(sums == k))
}

pOfTotalk(3, 8, 10000)
```
# Question 1
### Example: Dice Problem
Suppose three fair dice are rolled. We wish to find the approximate probability that the first die
shows fewer than 3 dots, given that the total number of dots for the 3 dice is more than 8, using
simulation.

```{r Dice conditional probability problem}

roll <- function(d) return(sample(1:6, d, replace = TRUE))

# r : number of dots on the first dice
# d : number of dice
# total : total sum from d dices
# nreps : number of repetitions
condionalDice <- function(r, d, total, nreps){
    count <- 0
    count2 <- 0
    for (i in 1:nreps){
        dice <- roll(d)
        if (sum(dice) > total){
            count <- count + 1
            if(dice[1] < r){
                count2 <- count2 + 1
            }
        }
    }
    return(count2/count)
}

condionalDice(3, 3, 8, 100000)
```

## ALOHA Problem
```{r, cache=TRUE}
sim <- function(p, q, nreps) {
    countx2eq2 <- 0
    countx2eq1 <- 0
    countx1eq2 <- 0
    
    
}
```